---
title: "Non parametric Manova"
output: html_document
---

```{r, include=FALSE}
options(width=200)
```

### `vegan` package

```{r}
library(vegan)
?adonis
```

```{r}
data(dune)
data(dune.env)
adonis(dune ~ Management*A1, data=dune.env, permutations=99)
```


Index de disimillllitud (default "bray"): "manhattan", "euclidean", "canberra", "bray", "kulczynski", "jaccard", "gower", "altGower", "morisita", "horn", "mountford", "raup" , "binomial", "chao", "cao" or "mahalanobis". Veure `?begdist` per la definició.

#### Referencies

Anderson, M.J. 2001. A new method for non-parametric multivariate analysis of variance. Austral Ecology, 26: 32–46.

------

### `ade4` package

```{r}
library(ade4)
?amova
```

```{r}
data(humDNAm)
amovahum <- amova(humDNAm$samples, sqrt(humDNAm$distances), humDNAm$structures)
amovahum
```


Quina forma tenen les dades?

* Mostra 

a data frame with haplotypes (or genotypes) as rows, populations as columns and abundance as entries

```{r}
str(humDNAm$samples)
```

* Distàncies

an object of class dist computed from Euclidean distance. If distances is null, equidistances are used.

```{r}
str(humDNAm$distances)
```

* Structures

a data frame containing, in the jth row and the kth column, the name of the group of level k to which the jth population belongs

```{r}
str(humDNAm$structures)
```

#### Referencies

Excoffier, L., Smouse, P.E. and Quattro, J.M. (1992) Analysis of molecular variance inferred from metric distances among DNA haplotypes: application to human mitochondrial DNA restriction data. Genetics, 131, 479–491.